---
title: "homework 2611"
author: "Anna Yermokhina"
format: html
---

## Homework 26.11

Aufgabe 1.

```{r}
library(tidyverse)

popA <- rnorm( 2000, mean=110, sd=20 )
popB <- rnorm( 2000, mean=120, sd=20 )

sample( popA, 10) -> sampleA
sample( popA, 10) -> sampleB

meanA = mean(sampleA)
meanB = mean(sampleB)

difference = meanB - meanA
difference
```

Aufgabe 2.

```{r}

diff_vector <-c(
replicate( 1000, {
popA <- rnorm( 2000, mean=110, sd= 20)
popB <- rnorm( 2000, mean=120, sd= 20)

sample( popA, 10) -> sampleA
sample( popB, 10) -> sampleB

meanA = mean(sampleA)
meanB = mean(sampleB)

diff = meanB - meanA
}))

hist(diff_vector, breaks = 30)
```

alternatives Histogramm um die Verteilungsdichte anzuzeigen

```{r}
diff_tbl <-
  tibble(
    difference = diff_vector) 

diff_tbl %>% 
ggplot() + 
  geom_histogram( aes( x = difference, y = after_stat( density)), bins = 20)
```

Aufgabe 3

für 10 Mäuse:

```{r}
B_better = sum( diff_vector > 0)
probability_B = sum( diff_vector > 0)/length( diff_vector)

A_better = sum( diff_vector < 0)
probability_A= sum( diff_vector  < 0)/length( diff_vector)

A_better
probability_A
B_better
probability_B
```

für 20 Mäuse:

```{r}

diff_vector <-c(
replicate( 1000, {
popA <- rnorm( 2000, mean=110, sd=20)
popB <- rnorm( 2000, mean= 120, sd=20)

sample( popA, 20) -> sampleA
sample( popB, 20) -> sampleB

meanA = mean(sampleA)
meanB = mean(sampleB)

diff = meanB - meanA
}))


B_better = sum( diff_vector > 0)
probability_B = sum( diff_vector > 0)/length( diff_vector)

A_better = sum( diff_vector < 0)
probability_A = sum( diff_vector  < 0)/length( diff_vector)

A_better
probability_A
B_better
probability_B
```

für 200 Mäuse:

```{r}
diff_vector <-c(
replicate( 1000, {
popA <- rnorm( 2000, mean=110, sd=20)
popB <- rnorm( 2000, mean= 120, sd=20)

sample( popA, 200) -> sampleA
sample( popB, 200) -> sampleB

meanA = mean(sampleA)
meanB = mean(sampleB)

diff = meanB - meanA
}))

diff_tbl <-
  tibble(
    difference = diff_vector) 

diff_tbl %>% 
ggplot() + 
  geom_histogram( aes( x = difference, y = after_stat( density)), bins= 30)

B_better = sum( diff_vector > 0)
probability_B = sum( diff_vector > 0)/length( diff_vector)

A_better = sum( diff_vector < 0)
probability_A = sum( diff_vector  < 0)/length( diff_vector)

A_better
probability_A
B_better
probability_B
```

Aufgabe 4.

By experimenting with sample size we could assume that collectying around 20 mice from each forest would be sufficient to determine if one of the populations had more nutrition.

However, if we were looking for a specific effect size, it would be great to conduct statistical power analysis.

Statistical power is 1-B, where B is the probability of the type 2 error (the probability not to reject H0 despite it being wrong).

We can use the following formula to find B:

B = P( Z \< Za - diff/sd( diff))

where diff is the difference between the two means, sd( diff) is the standard deviation of the difference and Za is the Z value for our significance level (1.96 for two-sided 95% confidence interval).

To determine sd( diff) we need to know the standard errors of the means: SEM = SD/sqrt( n)

The sd( diff) is a linear combination of the SEMs, because difference = meanB - meanA

sd( diff) = sqrt( SEMa\^2 + SEMb\^2)

Since the sample sizes and standard deviations of both groups are the same, we can simplify:

sd( diff) = sqrt( 2\* SEMa\^2) = sqrt( 2\* (SD/sqrt( n))\^2)) = SD\* sqrt( 2/n)

For example, for 10 mice:

B = P( Z \< Za - diff/sd( diff)) = P( Z \< 1.96 - 10/(SD\* sqrt( 2/n))) =

P( Z \< 1.96 - 10/(20\* sqrt( 2/10))) = P( Z \< 1.96 - 1.11) = P( Z \< 0.85) = 0.8

The power of the experiment with 10 mice from each population and 95% confidence is therefore

1 - 0.8 = 0.2 = 20%

For 20 mice:

B = P( Z \< Za - diff/sd( diff)) = P( Z \< 1.96 - 10/(SD\* sqrt( 2/n))) =

P( Z \< 1.96 - 10/(20\* sqrt( 2/20))) = P( Z \< 1.96 - 1.58) = P( Z \< 0.37) = 0.64

1 - 0.64 = 0.36

In the simulation with 20 mice we would be able to detect the 95% significant difference in about 36% of the cases.

This calculation can be reverted to find out the size of the sample required for a specific power.

For example, for 80% power:

B = 20% = 0.2 = P( Z \< Za - diff/sd( diff))

0.2 = P( Z \< Za - diff/sd( diff))

\- 0.85 = 1.96 - 10/(20\* sqrt( 2/n)))

-2.81 = - 10/(20\* sqrt( 2/n)))

2.81 = 10/(20\* sqrt( 2/n)))

5.62 = 1/sqrt( 2/n) = sqrt( n/2)

n/2 = 5.62\*5.62 = 31.6

n = 31.6\*2 = 63.1

Therefore, to achieve 80% statistical power for 95% significance level we would have to collect at least 64 mice from each forest.
